### 2019-04-11 initial/notes.tex
check in notes including tidal

Alternatively, consider Hamilton's equations applied to the Hamiltonian:
\begin{align}
    \pd{\phi}{t} = \pd{\mathcal{H}}{(\cos\theta)}
        &= \cos\theta - \eta\p*{\cos I + \sin I \cot \theta \cos \phi},\\
    \pd{(\cos \theta)}{t} = -\pd{\mathcal{H}}{\phi}
        &= +\eta \sin I \sin \theta \sin \phi.
\end{align}
This seems to produce the same trajectories as Kassandra's EOM\@!

@@

>>>>

Alternatively, consider Hamilton's equations applied to the Hamiltonian:
\begin{align}
    \pd{\phi}{t} = \pd{\mathcal{H}}{(\cos\theta)}
        &= \cos\theta - \eta\p*{\cos I + \sin I \cot \theta \cos \phi},\\
    \pd{(\cos \theta)}{t} = -\pd{\mathcal{H}}{\phi}
        &= +\eta \sin I \sin \theta \sin \phi.
\end{align}
This produces the same trajectories as the Cartesian EOM, so this is correct.
However, since $\pd{\phi}{t} \propto 1/\sin\theta$, this is not a desirable
system of equations to use, as they are very stiff near $\theta \approx 0$.

@@

\subsection{Tidal Dissipation}

We can add a tidal dissipation term; we write it in form
$\p*{\rd{\hat{s}}{t}}_{tide} = \epsilon \hat{s} \times \p*{\hat{l} \times
\hat{s}}$. Expanding, $\p*{\rd{\hat{s}}{t}}_{tide} = \epsilon \p*{\hat{z} -
s_z\hat{s}}$.
